---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.async

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.async`](https://sounkou-bioinfo.github.io/Rho/rho.async/) is the small
asynchronous contract at the base of [Rho](../../README.md). A `RhoTask`
represents one eventual value; a `RhoStream` represents an asynchronous
sequence. Both are open S7 protocols, so transports and worker systems can add
methods without changing callers.

## Tasks and streams

```{r compose}
library(rho.async)

tasks <- list(
  rho_task(20L),
  rho_then(rho_task(21L), function(value) value + 1L)
)
rho_await(rho_all(tasks), timeout = 1000)

stream <- rho_list_stream(c("start", "delta", "done"))
rho_stream_collect(stream, timeout = 1000)
```

Composition stays asynchronous. `rho_await()` is deliberately conspicuous and
is reserved for boundaries such as a CLI, a test, or an interactive request for
the finished value. Cancellation and timeout operations use the same task
contract; adapters own the mechanics of stopping their underlying handle.

## Interactive R

An interactive session may opt into a top-level task callback that drains work
already ready in the `later` loop after each completed R expression.

```{r task-callback, eval = FALSE}
bridge <- rho_task_callback_bridge()
bridge <- rho_await(rho_register_task_callback(bridge))

# Remove it explicitly when the session no longer needs the bridge.
bridge <- rho_await(rho_remove_task_callback(bridge))
```

The callback never waits for work and is not a scheduler. nanonext remains the
I/O substrate; the bridge only improves progress at interactive R safe points.

Continue with [`rho.http`](../rho.http/README.md), or see the
[`rho.async` reference](https://sounkou-bioinfo.github.io/Rho/rho.async/).
