---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache.path = "README_cache/"
)
```

# Rho

<!-- badges: start -->
[![R-CMD-check](https://github.com/sounkou-bioinfo/Rho/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sounkou-bioinfo/Rho/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

**An asynchronous agent runtime for R, inspired by
[Pi](https://github.com/badlogic/pi-mono).**

Rho expresses Pi's provider and agent architecture through S7 classes and open
generics. nanonext provides asynchronous I/O, and mirai provides worker
evaluation.

An effectful operation returns a task or stream. Waiting is an explicit edge
operation, so a CLI can block while a Shiny application, extension, or another
agent keeps composing work.

## An agent run

The deterministic provider is the executable specification for the same event
protocol used by live providers. `rho_prompt()` returns before the result is
collected; `rho_await()` marks the blocking boundary.

```{r agent}
library(rho.async)
library(rho.ai)
library(rho.agent)

agent <- rho_agent(
  provider = rho_faux_provider(),
  model = rho_model(provider = "faux", id = "faux")
)

run_task <- rho_prompt(agent, "hello from R")
rho_is_task(run_task)

run <- rho_await(run_task, timeout = 5000)
c(
  status = run@status,
  answer = run@messages[[2L]]@content[[1L]]@text
)
```

The transcript contains typed messages and content; the lifecycle is an ordered
sequence of typed events rather than callbacks hidden inside the provider.

```{r events}
vapply(run@events, function(event) event@type, character(1))
```

## OpenAI Codex with explicit credentials

Live authentication is a value passed to the provider catalog. Rho does not
search environment variables or process-global credential state. The importer
accepts a Pi or Codex auth file only when its path is supplied by the caller.

```{r codex-agent, cache = TRUE, cache.lazy = FALSE, cache.vars = "codex_example", cache.rebuild = isTRUE(getOption("rho.codex_readme_rebuild", FALSE))}
library(rho.coding)

credential_path <- getOption("rho.openai_codex_credential")
credential <- rho_load_openai_codex_credential(
  path = credential_path
) |>
  rho_await(timeout = 5000)

codex_provider <- rho_openai_codex_provider()
models <- rho_models(
  providers = list(codex_provider),
  credentials = rho_memory_credential_store(
    list(`openai-codex` = credential)
  )
)

codex_model <- rho_openai_codex_model("gpt-5.3-codex-spark")
codex_agent <- rho_agent(
  provider = models,
  model = codex_model,
  tools = list(rho_tool_r()),
  stream_options = list(reasoning_effort = "minimal")
)

codex_run <- rho_prompt(
  codex_agent,
  paste(
    "Call the r tool exactly once with code sum((1:100)^2).",
    "Then answer with only the integer result."
  )
) |>
  rho_await(timeout = 120000)

tool_result <- codex_run@tool_results[[1L]]
result <- tool_result@content[[1L]]@text

codex_example <- data.frame(
  model = codex_model@id,
  status = codex_run@status,
  tool = tool_result@tool_name,
  result = result
)
codex_example
```

Rebuild this example by supplying the credential file explicitly:

```bash
make rdm-codex CREDENTIAL=/absolute/path/to/auth.json
```

## The packages

Rho keeps transport, provider semantics, agent policy, and applications in
separate installable packages. Each package has a focused README and reference
site.

| package | role | documentation |
|---|---|---|
| **rho.async** | tasks, streams, cancellation, timeouts, and composition | [README](packages/rho.async/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.async/) |
| **rho.http** | typed HTTP requests, nanonext transport, and SSE decoding | [README](packages/rho.http/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.http/) |
| **rho.ai** | messages, models, capabilities, credentials, providers, and tools | [README](packages/rho.ai/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.ai/) |
| **rho.agent** | multi-turn execution, tool scheduling, queues, cancellation, and events | [README](packages/rho.agent/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.agent/) |
| **rho.ext** | asynchronous extension handlers and capability registration | [README](packages/rho.ext/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.ext/) |
| **rho.compute** | typed mirai expression and function-call tasks | [README](packages/rho.compute/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.compute/) |
| **rho.graphics** | declared graphics devices and hashed artifacts | [README](packages/rho.graphics/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.graphics/) |
| **rho.coding** | Bash, file, isolated-worker R, and explicit current-session R tools | [README](packages/rho.coding/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.coding/) |
| **rho.bio** | manifests, resolvers, receipts, and database-neutral SQL contracts | [README](packages/rho.bio/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.bio/) |
| **rho.duckdb** | DuckDB implementation of the asynchronous SQL contracts | [README](packages/rho.duckdb/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.duckdb/) |
| **rho.bio.agent** | bioinformatics tools registered through the extension API | [README](packages/rho.bio.agent/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.bio.agent/) |
| **rho.testkit** | bounded assertions for asynchronous tests | [README](packages/rho.testkit/README.md) · [reference](https://sounkou-bioinfo.github.io/Rho/rho.testkit/) |

Provider implementations with no independent dependency or ABI boundary live
in `rho.ai`. OpenAI Codex, GitHub Copilot, Z.ai, OpenAI, Anthropic, Ollama, and
the deterministic faux provider therefore share one typed provider surface
without a package per API. The [Pi parity ledger](docs/pi-parity.md) distinguishes
complete wire adapters from request translators whose normalized stream is not
yet complete. OpenAI, OpenAI Codex, GitHub Copilot, Z.ai, and Anthropic have
executable normalized-stream fixtures; the ledger records which external
account runs are still required before publication.

Bioinformatics remains downstream: `rho.bio`, `rho.duckdb`, and
`rho.bio.agent` consume the provider and agent substrate but do not define it.

## Install and develop

Rho targets R 4.4 or newer. While the repository is private, install from a
checkout:

```bash
git clone git@github.com:sounkou-bioinfo/Rho.git
cd Rho
make deps
make install
```

The authored API documentation is roxygen; the authored tests are R Markdown
files under each package's `inst/tinytest/rmd/` directory. Generated manuals,
namespaces, executable tests, and READMEs are reproducible from their sources.

```bash
make format       # Air
make rd           # roxygen2
make purl-tests   # Rmd tests -> executable tinytest files
make rdm          # rebuild package READMEs
make test
make check        # every package must report Status: OK
make public-ready # complete publication gate
```

The [Pi parity ledger](docs/pi-parity.md) records behavioral contracts and the
fixtures that verify them. Public release and addition to the
[sounkou-bioinfo R-universe](https://sounkou-bioinfo.r-universe.dev) follow
green package checks, live provider checks, documentation, and secret scanning.
The exact sequence is recorded in the [publishing guide](docs/releasing.md).

## License

MIT.
