---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.http

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.http`](https://sounkou-bioinfo.github.io/Rho/rho.http/) gives Rho one
typed HTTP boundary. Provider implementations build `RhoHttpRequest` values and
receive tasks or SSE streams; request encoding, TLS, and connection handling
remain in this package.

## TLS and SSE

```{r transport}
library(rho.http)

client <- rho_http_client()
inherits(client@tls, "tlsConfig")

events <- rho_sse_parse(paste0(
  "event: delta\n",
  "data: first\n\n",
  "data: done\n\n"
))
vapply(events, function(event) event@event, character(1))
vapply(events, function(event) event@data, character(1))
```

`rho_http_client()` uses nanonext's in-memory mbedTLS configuration. A caller
that requires peer authentication passes a configured `tlsConfig` value; the
package never searches operating-system certificate paths. SSE values feed
typed provider decoders in [`rho.ai`](../rho.ai/README.md).

See the [`rho.http` reference](https://sounkou-bioinfo.github.io/Rho/rho.http/)
and the underlying [`rho.async`](../rho.async/README.md) contract.
