---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.graphics

[`rho.graphics`](https://sounkou-bioinfo.github.io/Rho/rho.graphics/) makes R
graphics first-class asynchronous artifacts. A render declares its device and
returns dimensions, media type, digest, alternative text, and provenance; it
does not depend on whichever interactive device happens to be active.

## Render an artifact

```{r graphic}
library(rho.async)
library(rho.graphics)

path <- tempfile(fileext = ".png")
spec <- rho_base_graphic(
  plot(points, type = "b", xlab = "sample", ylab = "value"),
  data = list(points = c(2, 5, 3, 7))
)
artifact <- rho_render_graphic(
  spec,
  device = rho_png_device(width = 5, height = 3),
  path = path,
  alt = "Four connected sample values"
) |>
  rho_await(timeout = 10000)

list(
  format = artifact@format,
  media_type = artifact@media_type,
  alt = artifact@alt,
  digest = substr(artifact@sha256, 1L, 12L),
  bytes = unname(file.info(artifact@path)$size)
)
unlink(artifact@path)
```

PNG, SVG, and PDF are durable outputs. A recorded plot may support an ephemeral
preview, but it is not treated as archival provenance.

See the [`rho.graphics` reference](https://sounkou-bioinfo.github.io/Rho/rho.graphics/)
and its worker substrate, [`rho.compute`](../rho.compute/README.md).
