---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.bio.agent

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.bio.agent`](https://sounkou-bioinfo.github.io/Rho/rho.bio.agent/) exposes
declared bioinformatics resources to an agent through the ordinary
[`rho.ext`](../rho.ext/README.md) and [`rho.ai`](../rho.ai/README.md) tool
contracts. It is downstream of the provider and agent core.

## Register and call a bio tool

```{r bio-tool}
library(rho.async)
library(rho.ai)
library(rho.bio)
library(rho.ext)
library(rho.bio.agent)

registry <- rho_bio_registry()
rho_register_manifest(registry, rho_manifest(
  id = "example",
  version = "1.0.0",
  title = "Example manifest",
  description = "A registered manifest"
))

runtime <- rho_extension_runtime()
rho_register_bio_extension(rho_extension_api(runtime), registry)
tool <- get("bio_describe_manifest", runtime@state$tools)
result <- rho_execute_tool(
  tool,
  ToolCall("bio-1", "bio_describe_manifest", list()),
  context = NULL
) |>
  rho_await(timeout = 1000)

list(
  tool = tool@name,
  manifests = result@details$count
)
```

The extension translates registry facts into tool results; it does not make the
model the authority for those facts. Additional bio workflows can register
tools without changing `rho.agent`.

See the
[`rho.bio.agent` reference](https://sounkou-bioinfo.github.io/Rho/rho.bio.agent/)
and the underlying [`rho.bio`](../rho.bio/README.md) substrate.
