---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.http.httr2

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.http.httr2`](https://sounkou-bioinfo.github.io/Rho/rho.http.httr2/)
implements the [`rho.http`](https://sounkou-bioinfo.github.io/Rho/rho.http/)
client interface with httr2 connections owned by compute workers.

httr2 exposes a pull connection that parses response status, headers, transfer
framing, and body bytes. Opening that connection waits for the response head,
so the adapter does not run it on the calling R event loop. A worker owns the
connection and relays typed head, chunk, completion, and error values over a
private localhost NNG pair socket. The socket supplies backpressure, and closing
or cancelling the Rho stream cancels the worker.

```{r client}
library(rho.http.httr2)

client <- rho_httr2_http_client()
S7::S7_inherits(client, rho.http::RhoHttpClient)
s7contract::implements(client, rho.http::HttpClient)
rho.http::rho_http_client_close(client)
```

Providers receive only the `HttpClient` contract. Selecting this adapter does
not change request translation, SSE decoding, or normalized provider events.

See the [`rho.http.httr2` reference](https://sounkou-bioinfo.github.io/Rho/rho.http.httr2/reference/)
and the shared [`rho.http` contract](https://sounkou-bioinfo.github.io/Rho/rho.http/).
