---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.ai

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.ai`](https://sounkou-bioinfo.github.io/Rho/rho.ai/) defines the provider
surface shared by the rest of [Rho](https://sounkou-bioinfo.github.io/Rho/): typed messages and
content, model capabilities, tool contracts, credentials, provider operations,
and normalized assistant event streams.

## One stream protocol

```{r provider}
library(rho.async)
library(rho.ai)

provider <- rho_faux_provider()
model <- rho_model(provider = "faux", id = "faux")
context <- rho_context(messages = list(rho_user_message("hello")))

events <- rho_stream(provider, model, context) |>
  rho_stream_collect(timeout = 1000)
vapply(events, rho_assistant_event_type, character(1))
events[[length(events)]]@message@content[[1L]]@text
```

## Provider turns

`rho_stream()` is the normalized surface, not an HTTP operation. A model records
the turn transports it accepts, while a provider records the transports it
implements. Selection returns a typed strategy and its reason before the
provider opens an assistant-event stream.

```{r transport}
codex_provider <- rho_openai_codex_provider()@implementation
codex_model <- rho_openai_codex_model("gpt-5.3-codex-spark")
selection <- rho_select_provider_transport(
  codex_provider,
  codex_model,
  AutomaticTransport()
)

rho_transport_id(selection@transport)
selection@reason
```

SSE, WebSocket, cached WebSocket, and embedded execution are separate S7
classes. Provider packages implement `rho_provider_transports()` and
`rho_open_provider_transport()`; the broad `rho_stream()` method supplies the
common selection behavior. An embedded model can therefore return normalized
events directly, while a remote model can use HTTP, WebSocket, NNG, or another
implementation without changing the agent.

The faux provider is deterministic. OpenAI and OpenAI Codex reduce Responses
events to the same protocol; Z.ai reduces Chat Completions events; Anthropic
reduces Messages events. GitHub Copilot selects either the Responses or
Messages dialect from the model while retaining its own credentials and
endpoint headers. Capabilities are queried through values and generics instead
of provider-name conditionals:

```{r capabilities}
model <- rho_openai_codex_model("gpt-5.3-codex-spark")
list(
  text_input = rho_model_supports_input(model, "text"),
  image_input = rho_model_supports_input(model, "image"),
  thinking = rho_supported_thinking_levels(model)
)
```

## Hosted operations

Provider-hosted search is a semantic operation, not a local function tool.
The selected provider and model bind the operation before request translation:

```{r hosted-operation}
search_provider <- rho_openai_provider()
search_model <- rho_openai_model("gpt-5.4")
search_context <- rho_context(
  messages = list(rho_user_message("Find the current R release.")),
  operations = list(rho_web_search(
    domains = rho_web_search_allowed_domains("r-project.org")
  ))
)
search_plan <- rho_plan_operations(
  search_provider,
  search_model,
  search_context
)
binding <- search_plan@bindings[[1L]]

list(
  binding = S7::S7_class(binding)@name,
  reason = binding@reason
)
```

OpenAI and Anthropic bindings emit their own wire declarations and normalize
search calls, results, and citations as content. The agent executes only
`ToolCall` content from its local `ToolSpec` registry. An extension can provide
a narrower binding without changing either provider translator or the loop.

## Usage and cost

Token accounting is normalized before it reaches the agent. Ordinary input,
cache reads, and cache writes are disjoint; reasoning is an optional subset of
output. Pricing is an S7 generic, so the catalog supplies the common method and
specialized models can define different rules.

```{r usage}
usage <- rho_usage(
  input = 800,
  output = 120,
  cache_read = 200,
  reasoning = 40
)
priced <- rho_price_usage(model, usage)

c(
  tokens = priced@total,
  reasoning = priced@reasoning,
  input_cost = priced@cost@input,
  cache_read_cost = priced@cost@cache_read,
  total_cost = priced@cost@total
)
```

Credentials are explicit `RhoCredential` values resolved by a
`CredentialStore`. A request translator receives typed request auth; provider
code does not discover secrets through environment variables.

The memory store is process-scoped. The file store is selected with an explicit
path and persists successful login and refresh results as owner-readable JSON.
It uses the same `CredentialStore` methods, so provider auth code does not know
which storage implementation owns the credential.

GitHub Copilot uses device authorization and a short-lived session credential.
Z.ai keeps its Coding Plan endpoint, preserved-thinking policy, and streamed
tool-call policy in typed values. Both remain ordinary providers to the agent
loop.

Provider request configuration is typed. OpenAI and Anthropic request bodies
are reduced from S7 sections for tools, generation controls, caching, and
reasoning. Codex defaults are methods on `OpenAICodexResponsesModel`;
Anthropic thinking, temperature, cache, and tool-input behavior comes from a
typed model capability profile.

Continue with the [`rho.agent`](https://sounkou-bioinfo.github.io/Rho/rho.agent/) loop, or see the
[`rho.ai` reference](https://sounkou-bioinfo.github.io/Rho/rho.ai/reference/).
