---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.ai

[`rho.ai`](https://sounkou-bioinfo.github.io/Rho/rho.ai/) defines the provider
surface shared by the rest of [Rho](../../README.md): typed messages and
content, model capabilities, tool contracts, credentials, provider operations,
and normalized assistant event streams.

## One stream protocol

```{r provider}
library(rho.async)
library(rho.ai)

provider <- rho_faux_provider()
model <- rho_model(provider = "faux", id = "faux")
context <- rho_context(messages = list(rho_user_message("hello")))

events <- rho_stream(provider, model, context) |>
  rho_stream_collect(timeout = 1000)
vapply(events, rho_assistant_event_type, character(1))
events[[length(events)]]@message@content[[1L]]@text
```

The faux provider is deterministic, but it emits the same typed events as the
OpenAI, OpenAI Codex, Anthropic, and Ollama implementations. Capabilities are
queried through values and generics instead of provider-name conditionals:

```{r capabilities}
spark <- rho_openai_codex_spark()
list(
  text_input = rho_model_supports_input(spark, "text"),
  image_input = rho_model_supports_input(spark, "image"),
  thinking = rho_supported_thinking_levels(spark)
)
```

Credentials are explicit `RhoCredential` values resolved by a
`CredentialStore`. A request translator receives typed request auth; provider
code does not discover secrets through environment variables.

Continue with the [`rho.agent`](../rho.agent/README.md) loop, or see the
[`rho.ai` reference](https://sounkou-bioinfo.github.io/Rho/rho.ai/).
