---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.ai

[`rho.ai`](https://sounkou-bioinfo.github.io/Rho/rho.ai/) defines the provider
surface shared by the rest of [Rho](../../README.md): typed messages and
content, model capabilities, tool contracts, credentials, provider operations,
and normalized assistant event streams.

## One stream protocol

```{r provider}
library(rho.async)
library(rho.ai)

provider <- rho_faux_provider()
model <- rho_model(provider = "faux", id = "faux")
context <- rho_context(messages = list(rho_user_message("hello")))

events <- rho_stream(provider, model, context) |>
  rho_stream_collect(timeout = 1000)
vapply(events, rho_assistant_event_type, character(1))
events[[length(events)]]@message@content[[1L]]@text
```

The faux provider is deterministic. OpenAI, OpenAI Codex, and GitHub Copilot
reduce Responses events to the same protocol; Z.ai reduces Chat Completions
events to it. Capabilities are queried through values and generics instead of
provider-name conditionals:

```{r capabilities}
spark <- rho_openai_codex_spark()
list(
  text_input = rho_model_supports_input(spark, "text"),
  image_input = rho_model_supports_input(spark, "image"),
  thinking = rho_supported_thinking_levels(spark)
)
```

## Usage and cost

Token accounting is normalized before it reaches the agent. Ordinary input,
cache reads, and cache writes are disjoint; reasoning is an optional subset of
output. Pricing is an S7 generic, so the catalog supplies the common method and
specialized models can define different rules.

```{r usage}
usage <- rho_usage(
  input = 800,
  output = 120,
  cache_read = 200,
  reasoning = 40
)
priced <- rho_price_usage(spark, usage)

c(
  tokens = priced@total,
  reasoning = priced@reasoning,
  input_cost = priced@cost@input,
  cache_read_cost = priced@cost@cache_read,
  total_cost = priced@cost@total
)
```

Credentials are explicit `RhoCredential` values resolved by a
`CredentialStore`. A request translator receives typed request auth; provider
code does not discover secrets through environment variables.

GitHub Copilot uses device authorization and a short-lived session credential.
Z.ai keeps its Coding Plan endpoint, preserved-thinking policy, and streamed
tool-call policy in typed values. Both remain ordinary providers to the agent
loop.

OpenAI request configuration is also typed. A request body is reduced from S7
sections for tools, generation controls, cache affinity, and reasoning. Codex
defaults are methods on `OpenAICodexResponsesModel`, not conditionals in the
standard OpenAI builder.

Continue with the [`rho.agent`](../rho.agent/README.md) loop, or see the
[`rho.ai` reference](https://sounkou-bioinfo.github.io/Rho/rho.ai/).
