---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.coding

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.coding`](https://sounkou-bioinfo.github.io/Rho/rho.coding/) supplies
coding tools with explicit execution semantics: file operations, Bash with a
typed cross-platform resolution, isolated R evaluation in mirai, and opt-in
evaluation in a caller-supplied current-session environment.

## Isolated R evaluation

```{r r-tool}
library(rho.async)
library(rho.ai)
library(rho.agent)
library(rho.coding)

r_tool <- rho_tool_r()
result <- rho_execute_tool(
  r_tool,
  ToolCall(
    id = "readme-r-1",
    name = "r",
    arguments = list(code = "sum((1:6)^2)")
  ),
  context = NULL
) |>
  rho_await(timeout = 10000)

data.frame(
  value = result@details$value,
  may_overlap = S7::S7_inherits(r_tool@overlap, ToolMayOverlap)
)
```

The ordinary R tool is isolated and may overlap with another call. A
`RhoCurrentSessionREvaluator` instead receives an explicit environment and
requires exclusive scheduling. `RhoRExpression` is also a `RhoOperation`, and
the chosen evaluator is recorded in a `RhoREvaluationBinding`. A remote NNG
evaluator therefore adds evaluator methods rather than another agent execution
path. Bash follows the same discipline: on Windows it resolves a real Bash
implementation rather than translating model-generated Bash into another shell
language, and on Unix it reports a typed unavailable value when Bash is absent.

## Session replay

The coding host may persist the agent journal as locked JSONL without making a
path part of the agent contract. The codec derives its supported document types
from the S7 classes reachable from session entries and messages.

```{r session-journal}
path <- tempfile(fileext = ".jsonl")
journal <- rho_jsonl_session_journal(path)

writer <- rho_agent(
  rho_faux_provider(),
  rho_model("faux", "faux"),
  journal = journal
)
run <- rho_prompt(writer, "remember this turn") |>
  rho_await(timeout = 10000)

reader <- rho_agent(
  rho_faux_provider(),
  rho_model("faux", "faux"),
  journal = rho_jsonl_session_journal(path)
)
reader <- rho_sync_session(reader) |>
  rho_await(timeout = 10000)

snapshot <- rho_session_snapshot(reader@journal) |>
  rho_await(timeout = 10000)

data.frame(
  committed_entries = snapshot@position,
  restored_messages = length(rho_state_messages(reader))
)

unlink(c(path, paste0(path, ".lock")))
```

The native format is versioned Rho JSONL, not Pi session JSONL. Pi import or
export can be added as another codec once session identity and branch lineage
are exercised by Rho consumers.

See the [`rho.coding` reference](https://sounkou-bioinfo.github.io/Rho/rho.coding/reference/)
and its worker substrate, [`rho.compute`](https://sounkou-bioinfo.github.io/Rho/rho.compute/).
