---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.coding

[`rho.coding`](https://sounkou-bioinfo.github.io/Rho/rho.coding/) supplies
coding tools with explicit execution semantics: file operations, Bash with a
typed cross-platform resolution, isolated R evaluation in mirai, and opt-in
evaluation in a caller-supplied current-session environment.

## Isolated R evaluation

```{r r-tool}
library(rho.async)
library(rho.ai)
library(rho.coding)

r_tool <- rho_tool_r()
result <- rho_execute_tool(
  r_tool,
  ToolCall(
    id = "readme-r-1",
    name = "r",
    arguments = list(code = "sum((1:6)^2)")
  )
) |>
  rho_await(timeout = 10000)

list(
  value = result@details$value,
  may_overlap = S7::S7_inherits(r_tool@overlap, ToolMayOverlap)
)
```

The ordinary R tool is isolated and may overlap with another call. A
`RhoCurrentSessionREvaluator` instead receives an explicit environment and
requires exclusive scheduling. Bash follows the same discipline: on Windows it
resolves a real Bash implementation rather than translating model-generated
Bash into another shell language.

See the [`rho.coding` reference](https://sounkou-bioinfo.github.io/Rho/rho.coding/)
and its worker substrate, [`rho.compute`](../rho.compute/README.md).
