---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.ext

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.ext`](https://sounkou-bioinfo.github.io/Rho/rho.ext/) is the open
extension runtime for [Rho](https://sounkou-bioinfo.github.io/Rho/). Extensions register handlers,
tools, commands, and providers through an explicit API object; they do not
patch package globals.

## An asynchronous handler chain

```{r extension}
library(rho.async)
library(rho.ext)

runtime <- rho_extension_runtime()
api <- rho_extension_api(runtime, source = "README")
rho_on(api, "prompt", function(event, context) {
  rho_task(list(text = toupper(event$text)))
})

results <- rho_dispatch_event(
  runtime,
  list(type = "prompt", text = "compose me")
) |>
  rho_await(timeout = 1000)
results[[1L]]
```

The dispatcher normalizes plain handler values and tasks. This
lets an extension perform I/O, ask for permission, or consult policy without
turning the agent loop into a synchronous callback system.

See the [`rho.ext` reference](https://sounkou-bioinfo.github.io/Rho/rho.ext/reference/),
the upstream [`rho.agent`](https://sounkou-bioinfo.github.io/Rho/rho.agent/) lifecycle, and the downstream
[`rho.bio.agent`](https://sounkou-bioinfo.github.io/Rho/rho.bio.agent/) application.
