---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.testkit

[`rho.testkit`](https://sounkou-bioinfo.github.io/Rho/rho.testkit/) supplies
bounded tinytest assertions for Rho tasks and streams. Every assertion has an
explicit timeout, so a broken asynchronous path fails instead of hanging the
test process.

## Bounded assertions (runs at render time)

```{r assertions}
library(rho.async)
library(rho.testkit)

value <- expect_resolves(
  rho_then(rho_task(20L), function(x) x + 22L),
  expected = 42L,
  timeout = 1000
)
items <- expect_stream_length(
  rho_list_stream(letters[1:3]),
  n = 3L,
  timeout = 1000
)
list(value = value, items = items)
```

The package also provides a deterministic faux-provider agent fixture. Authored
tests remain R Markdown documents; the monorepo derives executable tinytest
files and checks that generated files are current.

See the [`rho.testkit` reference](https://sounkou-bioinfo.github.io/Rho/rho.testkit/)
and the base [`rho.async`](../rho.async/README.md) contracts.
