---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# rho.testkit

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

[`rho.testkit`](https://rgenomicsetl.github.io/Rho/rho.testkit/) supplies
bounded tinytest assertions for Rho tasks and streams. Every assertion has an
explicit timeout, so a broken asynchronous path fails instead of hanging the
test process.

## Bounded assertions

```{r assertions}
library(rho.async)
library(rho.testkit)

value <- expect_resolves(
  rho_then(rho_task(20L), function(x) x + 22L),
  expected = 42L,
  timeout = 1000
)
items <- expect_stream_length(
  rho_list_stream(letters[1:3]),
  n = 3L,
  timeout = 1000
)
list(value = value, items = items)
```

The package also provides a deterministic faux-provider agent fixture. Authored
tests remain R Markdown documents; the monorepo derives executable tinytest
files and checks that generated files are current.

See the [`rho.testkit` reference](https://rgenomicsetl.github.io/Rho/rho.testkit/reference/)
and the base [`rho.async`](https://rgenomicsetl.github.io/Rho/rho.async/) contracts.
